---
title: "Organizing ENIGH Household Data for 2016"
author: "Mauricio Hernandez"
date: "3/10/2021"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Clear memory
rm(list=ls())

#Set WD
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
print( getwd() )

#Load Packages
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(pacman, party, psych, rio, tidyverse, plyr, knitr, bit64, 
                 forcats, stargazer, summarytools)

opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Reading ENIGH microdata CSV files
```{r include= TRUE, echo = FALSE}
path.data <- "https://mauricioh2.com/disreu/inegi/enigh/"


df_hshd <-read_csv(paste0(path.data, "2016_enigh_hogares.csv"), col_names = TRUE,
                   col_types=c("folioviv"="c", "foliohog"="c", 'anio_carret'='i', 
                               'anio_otro'='i', 'trab_come' ='i', 'huesp_come'='i', 
                               'anio_auto' = 'i', 'anio_van' = 'i', 
                               'anio_pickup' = 'i', 'anio_moto' = 'i', 
                               'anio_bici' = 'i', 'anio_trici' = 'i', 
                               'anio_canoa' = 'i', 'anio_ester' = 'i', 
                               'anio_grab' = 'i', 'anio_radio' = 'i', 
                               'anio_tva' = 'i', 'anio_tvd' = 'i', 
                               'anio_dvd' = 'i', 'anio_video' = 'i', 
                               'anio_licua' = 'i', 'anio_tosta' = 'i', 
                               'anio_micro' = 'i', 'anio_refri' = 'i', 
                               'anio_estuf' = 'i', 'anio_lavad' = 'i', 
                               'anio_planc' = 'i', 'anio_maqui' = 'i', 
                               'anio_venti' = 'i', 'anio_aspir' = 'i', 
                               'anio_compu' = 'i', 'anio_impre' = 'i',
                               'nr_viv' = 'i',
                               'anio_juego'='i'), na ='&')
```


## Organizing Household data
Attribute folioviv is the identifier of the dwelling. It is made up of two digits with the federal entity code, one digit with the scope (urban, code different from 6; rural, code 6), four digits of the consecutive number of the upm, one digit of the ten of survey and two digits with a consecutive number for the selected dwelling.

```{r}
df_hshd <- as_tibble(df_hshd) %>%
  select("folioviv", "foliohog", "huespedes", "huesp_come", "num_trab_d", "trab_come", "telefono", "celular", "tv_paga", "conex_inte", "num_auto", "anio_auto", "num_van", "anio_van", "num_pickup", "anio_pickup", "num_moto", "anio_moto", "num_bici", "anio_bici", "num_trici", "anio_trici", "num_carret", "anio_carret", "num_canoa", "anio_canoa", "num_otro", "anio_otro", "num_ester", "anio_ester", "num_grab", "anio_grab", "num_radio", "anio_radio", "num_tva", "anio_tva", "num_tvd", "anio_tvd", "num_dvd", "anio_dvd", "num_video", "anio_video", "num_licua", "anio_licua", "num_tosta", "anio_tosta", "num_micro", "anio_micro", "num_refri", "anio_refri", "num_estuf", "anio_estuf", "num_lavad", "anio_lavad", "num_planc", "anio_planc", "num_maqui", "anio_maqui", "num_venti", "anio_venti", "num_aspir", "anio_aspir", "num_compu", "anio_compu", "num_impre", "anio_impre", "num_juego", "anio_juego", "esc_radio", "er_aparato", "er_celular", "er_compu", "er_aplicac", "er_tv", "er_otro", "recib_tvd")


df_hshd$state_code <- substr(df_hshd$folioviv, 1, 2) %>% strtoi(base = 10L)
df_hshd$urbanization <- substr(df_hshd$folioviv, 3, 3) %>% strtoi(base = 10L)
df_hshd$upm <- substr(df_hshd$folioviv, 4, 7) %>% strtoi(base = 10L)
df_hshd$ten_surv <- substr(df_hshd$folioviv, 8, 8) %>% strtoi(base = 10L)
df_hshd$consecutive_dwell <- substr(df_hshd$folioviv, 9, 10) %>% strtoi(base = 0L)

# Create/convert the new variable state into a factor, and assign name of the state as label
df_hshd$state <- factor(df_hshd$state_code, 
                     levels = c("1", "2", "3", "4", "5", "6", "7", 
                                "8", "9", "10","11", "12", "13", "14", 
                                "15", "16", "17", "18", "19", "20", 
                                "21", "22", "23", "24", "25", "26", 
                                "27", "28", "29", "30", "31", "32"),
                     labels = c("Aguascalientes", "Baja California", 
                                "Baja California Sur", "Campeche", 
                                "Coahuila", "Colima", "Chiapas", 
                                "Chihuahua", "Mexico City", 
                                "Durango", "Guanajuato", "Guerrero", 
                                "Hidalgo", "Jalisco", "Mexico", 
                                "Michoacan", "Morelos", "Nayarit", 
                                "Nuevo Leon", "Oaxaca", "Puebla", 
                                "Queretaro", "Quintana Roo", 
                                "San Luis Potosi", "Sinaloa", 
                                "Sonora", "Tabasco", "Tamaulipas", 
                                "Tlaxcala", "Veracruz", "Yucatan", 
                                "Zacatecas"))
```


```{r, include= FALSE}
str(df_hshd)
```


```{r}
# if urban, urbanization code different from 6; rural, code 6
# In new attribute rural, 1 if rural, 0 = urban
df_hshd$rural <- NA
df_hshd <- within(df_hshd, rural[urbanization!=6] <- 0)
df_hshd <- within(df_hshd, rural[urbanization==6] <- 1)                 
count(df_hshd, 'rural')

# Create household id, this will be the key attribute in the SQL table
df_hshd$id_household <- paste0(df_hshd$folioviv, df_hshd$foliohog)
head(df_hshd$id_household)

# Changing yes/no questions, in the original data of the survey yes =1, no = 2, 
# here yes = 1, no = 0
df_hshd$telefono[df_hshd$telefono == 2] <- 0
df_hshd$celular[df_hshd$celular == 2] <- 0
df_hshd$tv_paga[df_hshd$tv_paga == 2] <- 0
df_hshd$conex_inte[df_hshd$conex_inte == 2] <- 0
df_hshd$esc_radio[df_hshd$esc_radio == 2] <- 0

df_hshd$esc_radio[df_hshd$esc_radio == 2] <- 0

#df_hshd$anio_auto  [df_hshd$anio_auto  >= 95] <- NA
#df_hshd$anio_van   [df_hshd$anio_van   >= 95] <- NA
#df_hshd$anio_pickup[df_hshd$anio_pickup>= 95] <- NA
#df_hshd$anio_moto  [df_hshd$anio_moto  >= 95] <- NA
#df_hshd$anio_bici  [df_hshd$anio_bici  >= 95] <- NA
#df_hshd$anio_canoa [df_hshd$anio_canoa >= 95] <- NA
#df_hshd$anio_otro  [df_hshd$anio_otro  >= 95] <- NA
#df_hshd$num_ester  [df_hshd$num_ester  >= 95] <- NA
#df_hshd$anio_ester [df_hshd$anio_ester >= 95] <- NA
#df_hshd$anio_grab  [df_hshd$anio_grab  >= 95] <- NA
#df_hshd$anio_radio [df_hshd$anio_radio >= 95] <- NA
#df_hshd$anio_tva   [df_hshd$anio_tva   >= 95] <- NA
#df_hshd$anio_tvd   [df_hshd$anio_tvd   >= 95] <- NA
#df_hshd$anio_dvd   [df_hshd$anio_dvd   >= 95] <- NA
#df_hshd$anio_video [df_hshd$anio_video >= 95] <- NA
#df_hshd$anio_licua [df_hshd$anio_licua >= 95] <- NA
#df_hshd$anio_tosta [df_hshd$anio_tosta >= 95] <- NA
#df_hshd$anio_micro [df_hshd$anio_micro >= 95] <- NA
#df_hshd$anio_refri [df_hshd$anio_refri >= 95] <- NA
#df_hshd$anio_estuf [df_hshd$anio_estuf >= 95] <- NA
#df_hshd$anio_lavad [df_hshd$anio_lavad >= 95] <- NA
#df_hshd$anio_planc [df_hshd$anio_planc >= 95] <- NA
#df_hshd$anio_maqui [df_hshd$anio_maqui >= 95] <- NA
#df_hshd$anio_venti [df_hshd$anio_venti >= 95] <- NA
#df_hshd$anio_aspir [df_hshd$anio_aspir >= 95] <- NA
#df_hshd$anio_compu [df_hshd$anio_compu >= 95] <- NA
#df_hshd$anio_impre [df_hshd$anio_impre >= 95] <- NA
#df_hshd$anio_juego [df_hshd$anio_juego >= 95] <- NA
```


```{r}
#summary(df_hshd, digits = 1)
#Ref: https://www.jakeruss.com/cheatsheets/stargazer/
#stargazer(df_hshd, summary.logical = FALSE)
descr(df_hshd, stats = c("mean", "sd", "min", "med", "max", "n.valid", "pct.valid"), 
      transpose = TRUE)
```


```{r}
#These values are stored in the dwelling table
df_hshd <- subset(df_hshd, select= c(-state_code, -urbanization, -upm, -ten_surv, -consecutive_dwell))

write.csv(df_hshd, file = 'outputs/2016_enigh_households.csv', na = "NULL", 
          fileEncoding="UTF-8", row.names=FALSE)
```