---
title: "Organizing ENIGH Dwelling Data"
author: "Mauricio Hernandez"
date: "2/21/2021"
output: html_notebook
---

Follow convention https://www.r-bloggers.com/2019/01/%F0%9F%96%8A-r-coding-style-guide

Take data from INEGI 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Clear memory
rm(list=ls())

#Set WD
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
print( getwd() )

#Load Packages
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(pacman, party, psych, rio, tidyverse, plyr, knitr, bit64, 
                 forcats, stargazer, summarytools)

opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Reading ENIGH microdata CSV files
```{r include= TRUE, echo = FALSE}
path.data <- "https://mauricioh2.com/disreu/inegi/enigh/"


#df_dwell <-read_csv(paste0(path.data, "2018_enigh_viviendas.csv"), col_names = TRUE,
#                    col_types=c("folioviv"="c", 'estufa_chi'='i'), na ='&')

df_dwell <-read_csv(paste0(path.data, "2018_enigh_viviendas.csv"), col_names = TRUE,
                   col_types=c("folioviv"="c", "estufa_chi"= 'i', 'mat_techos' ='i', 
                               'num_dueno1' ='i', 'num_dueno2' ='i'), na ='&')

```

```{r}
head(df_dwell)
```


## Organizing Household data
# attribute folioviv is the identifier of the dwelling. It is made up of two digits with the federal entity code, one digit with the scope (urban, code different from 6; rural, code 6), four digits of the consecutive number of the upm, one digit of the ten of survey and two digits with a consecutive number for the selected dwelling.

```{r}
df_dwell <- as_tibble(df_dwell) %>%
  select("folioviv", "tipo_viv", "mat_pared", "mat_techos", "mat_pisos", "antiguedad", 
         "antigua_ne", "cocina","cocina_dor", "cuart_dorm", "num_cuarto", "disp_agua", 
         "dotac_agua", "excusado", "uso_compar", "sanit_agua", "biodigest", "bano_comp", 
         "bano_excus", "bano_regad", "drenaje", "disp_elect", "focos_inca", "focos_ahor",
         "combustible", "estufa_chi", "tenencia", "renta", "estim_pago", "pago_viv", 
         "tipo_adqui", "viv_usada", "tipo_finan","num_dueno1", "lavadero", "fregadero", 
         "regadera", "tinaco_azo", "cisterna", "pileta", "calent_sol", "calent_gas", 
         "medidor_luz", "bomba_agua", "tanque_gas", "aire_acond", "calefacc", "tot_resid", 
         "tot_hom", "tot_muj", "tot_hog", "ubica_geo", "tam_loc", "est_socio", "est_dis", 
         "upm", "factor", "procaptar")

df_dwell$folioviv_state_code <- substr(df_dwell$folioviv, 1, 2) %>% strtoi(base = 10L)
df_dwell$folioviv_urban <- substr(df_dwell$folioviv, 3, 3) %>% strtoi(base = 10L)
df_dwell$folioviv_upm <- substr(df_dwell$folioviv, 4, 7) %>% strtoi(base = 10L)
df_dwell$folioviv_ten_surv <- substr(df_dwell$folioviv, 8, 8) %>% strtoi(base = 10L)
df_dwell$folioviv_consec_dwell <- substr(df_dwell$folioviv, 9, 10) %>% strtoi(base = 0L)

# Create/convert the new variable state into a factor, and assign name of the state as label
df_dwell$state <- factor(df_dwell$folioviv_state_code, 
                     levels = c("1", "2", "3", "4", "5", "6", "7", 
                                "8", "9", "10","11", "12", "13", "14", 
                                "15", "16", "17", "18", "19", "20", 
                                "21", "22", "23", "24", "25", "26", 
                                "27", "28", "29", "30", "31", "32"),
                     labels = c("Aguascalientes", "Baja California", 
                                "Baja California Sur", "Campeche", 
                                "Coahuila", "Colima", "Chiapas", 
                                "Chihuahua", "Mexico City", 
                                "Durango", "Guanajuato", "Guerrero", 
                                "Hidalgo", "Jalisco", "Mexico", 
                                "Michoacan", "Morelos", "Nayarit", 
                                "Nuevo Leon", "Oaxaca", "Puebla", 
                                "Queretaro", "Quintana Roo", 
                                "San Luis Potosi", "Sinaloa", 
                                "Sonora", "Tabasco", "Tamaulipas", 
                                "Tlaxcala", "Veracruz", "Yucatan", 
                                "Zacatecas"))
```



Contiene la ubicación geográfica de la vivienda. 
Los dos primeros dígitos representan la clave  de la entidad y los siguientes tres la clave del 
municipio. Éstas corresponden al Catálogo de claves de entidades federativas, municipios y localidades, que está disponible en el sitio del INEGI.

Contains the geographic location of the dwelling.
The first two digits represent the entity's key and the next three the entity's key.
municipality. These correspond to the Catalog of codes for states, municipalities and localities, which is available on INEGI's website.


```{r}
df_dwell$ubica_geo_state <- substr(df_dwell$ubica_geo, 1, 2) %>% strtoi(base = 10L)
df_dwell$ubica_geo_munic <- substr(df_dwell$ubica_geo, 3, 5) %>% strtoi(base = 10L)

#head(df_dwell, c('ubica_geo', 'ubica_geo_state', 'ubica_geo_munic'))

tail(subset(df_dwell, select=c('ubica_geo', 'ubica_geo_state', 'ubica_geo_munic')))
```

```{r}
str(df_dwell)
```


```{r}
# if urban, urbanization code different from 6; rural, code 6
# In new attribute rural, 1 if rural, 0 = urban
df_dwell$rural <- NA
df_dwell <- within(df_dwell, rural[folioviv_urban!=6] <- 0)
df_dwell <- within(df_dwell, rural[folioviv_urban==6] <- 1)                 
count(df_dwell, 'rural')

# Changing yes/no questions, in the original data of the survey yes =1, no = 2, 
# here yes = 1, no = 0
df_dwell$cocina[df_dwell$cocina == 2] <- 0
df_dwell$cocina_dor[df_dwell$cocina_dor == 2] <- 0
df_dwell$excusado[df_dwell$excusado == 2] <- 0
df_dwell$uso_compar[df_dwell$uso_compar == 2] <- 0
df_dwell$biodigest[df_dwell$biodigest == 2] <- 0
df_dwell$estufa_chi[df_dwell$estufa_chi == 2] <- 0

df_dwell$viv_usada[df_dwell$viv_usada == 2] <- 0
df_dwell$lavadero[df_dwell$lavadero == 2] <- 0
df_dwell$fregadero[df_dwell$fregadero == 2] <- 0
df_dwell$regadera[df_dwell$regadera == 2] <- 0
df_dwell$tinaco_azo[df_dwell$tinaco_azo == 2] <- 0
df_dwell$cisterna[df_dwell$cisterna == 2] <- 0
df_dwell$pileta[df_dwell$pileta == 2] <- 0
df_dwell$calent_sol[df_dwell$calent_sol == 2] <- 0
df_dwell$calent_gas[df_dwell$calent_gas == 2] <- 0

df_dwell$medidor_luz[df_dwell$medidor_luz == 2] <- 0
df_dwell$bomba_agua[df_dwell$bomba_agua == 2] <- 0
df_dwell$tanque_gas[df_dwell$tanque_gas == 2] <- 0
df_dwell$aire_acond[df_dwell$aire_acond == 2] <- 0
df_dwell$calefacc[df_dwell$calefacc == 2] <- 0
```

```{r}
head(subset(df_dwell, select=c(folioviv, num_dueno1, ubica_geo, est_dis, upm, state)))
```

```{r}
descr(df_dwell, stats = c("mean", "sd", "min", "med", "max", "n.valid", "pct.valid"), 
      transpose = TRUE)
```


```{r}
write.csv(df_dwell, file = 'outputs/2018_enigh_dwellings.csv', na = "NULL", 
          fileEncoding="UTF-8", row.names=FALSE)
```