---
title: "ENIGH Data Analysis"
author: "Mauricio Hernandez"
date: "2/22/2021"
subtitle: Household's Energy Related Expenses 
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE}
#Remove all objects from current workspace and call garbage collector
rm(list=ls())
gc()

#Set WD
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
```

```{r sh, include=FALSE}
source("./scripts/setup_functions.R")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
	             fig.width = 8, 
	             collapse = TRUE,
               fig.path = 'figs/')
```

```{r knitr_init, include=FALSE}
knitr::knit_hooks$set(output = function(x, options){
  if(!is.null(options$max_height)){
    paste('<pre style = "max-height:',
          options$max_height, 
          ';float: left; width: 910px; overflow-y: auto;">',
          x,
          "</pre>",
          sep = "")
  } else{
    x
  }
})


#Set WD
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
print( getwd() )

#Load Packages
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(pacman, party, psych, rio, tidyverse, knitr, bit64, plotly, 
                 forcats, stargazer, summarytools, survey, scales)

opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


```{r, include=FALSE}
#Generates the css needed by summarytools
st_css()
```

## Reading ENIGH microdata (CSV files)
```{r include = TRUE, echo = FALSE,  results='hide'}
path_data <- "https://mauricioh2.com/disreu/inegi/enigh/"

df_dwell <-read_csv(paste0(path_data, "2018_enigh_viviendas.csv"), col_names = TRUE,
                   col_types=c("folioviv"="c", "estufa_chi"= 'i', 'mat_techos' ='i', 
                               'num_dueno1' ='i', 'num_dueno2' ='i'), na ='&')

df_hsd_energy <-read_csv('./outputs/2018_enigh_hsd_energy_exp.csv', 
                         col_names = TRUE, progress = TRUE, na ='NULL',
                         col_types=cols_only("folioviv"='c', "foliohog"='c', 
                                             "clave"='c',
                                             "tipo_gasto"='c', "mes_dia"='c', 
                                             "forma_pag1"='i', "forma_pag2"='i', 
                                             "forma_pag3"='i', "lugar_comp"='i',
                                             "orga_inst"='i', "frecuencia"='i', 
                                             "fecha_adqu"='c', "fecha_pago"='c', 
                                             "cantidad"='d', "gasto"='d',
                                             "pago_mp"='d', "costo"='d', 
                                             "inmujer"='d',"inst_1"='i', 
                                             "inst_2"='i', "num_meses"='i',
                                             "num_pagos"='i', "ultim_pago"='c', 
                                             "gasto_tri"='d',"gasto_nm"='d', 
                                             "gas_nm_tri"='d', "imujer_tri"='d'),)
```


```{r include = TRUE, echo = FALSE,  results='hide'}
df_person_energy <-read_csv('./outputs/2018_enigh_psn_energy_exp.csv',col_names = TRUE,
                   col_types=c("folioviv"="c", 'inst' ='i', 'colegia' = 'd', 
                               'material' = 'd', 'inscrip' = 'd', 'gasto_tri'= 'd'), 
                   na ='NULL',)
```

## Check Energy Expenses

Check all the energy expenses in the households expenses data set and people expenses data sets are the same than the ones reported in the households condensed data set.

df_hsd_condens$energia is equal to the sum of gastoshogar.gasto_tri if clave is equal to G009-G016, R001, R003 plus 
gastospersona.gasto_tri if clave is equal to G009-G016, R001, R003.


|Code (clave) |    Description         |
| ----------- |:----------------------:|
|G009    	    |Liquefied petroleum gas |
|G010    	    |Petroleum               |
|G011    	    |Diesel                  |
|G012    	    |Carbon                  |
|G013    	    |Firewood                |
|G014    	    |Fuel to heat            |
|G015    	    |Candles                 |
|G016    	    |Other fuels             |
|R001    	    |Electricity             |
|R003    	    |Natural Gas			       |


```{r}
#df_person_energy
# Create household id, this will be the key attribute in the SQL table
df_hsd_energy$id_household <- paste0(df_hsd_energy$folioviv, df_hsd_energy$foliohog)
df_person_energy$id_household <- paste0(df_person_energy$folioviv, df_person_energy$foliohog)
```


Grouping data sets by household id and expense id (clave)
```{r, results='hide'}
df_hsd_energy_by_hsd <- df_hsd_energy %>%
  group_by(id_household, clave) %>% replace(is.na(.), 0) %>% 
  summarise(gasto_tri = sum(gasto_tri))

df_psn_energy_by_hsd <- df_person_energy %>%
  group_by(id_household, clave) %>% replace(is.na(.), 0) %>% 
  summarise(gasto_tri = sum(gasto_tri))

head(df_hsd_energy_by_hsd)
head(df_psn_energy_by_hsd)
```
Merging data sets by id_household using energy expenses related data
```{r}
df_energy_by_hsd <- merge(df_hsd_energy_by_hsd, df_psn_energy_by_hsd, 
                          by="id_household", all.y = TRUE, all.x = TRUE) %>% 
  replace(is.na(.), 0)
```

Reshaping data
```{r}
df_hsd_energy_by_hsd <- as.data.frame(df_hsd_energy_by_hsd)

df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G009'] <- 'LPG'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G010'] <- 'Petroleum'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G011'] <- 'Diesel'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G012'] <- 'Carbon'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G013'] <- 'Firewood'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G014'] <- 'Fuel_to_heat'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G015'] <- 'Candles'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='G016'] <- 'Other'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='R001'] <- 'Electricity'
df_hsd_energy_by_hsd$clave[df_hsd_energy_by_hsd$clave=='R003'] <- 'NG'

#df_hsd_energy_by_hsd$clave <-as.factor(df_hsd_energy_by_hsd$clave)
df_hsd_by_hsd_wd <- reshape(df_hsd_energy_by_hsd, v.names = 'gasto_tri',
                            timevar='clave', idvar="id_household", sep = "_",
                            direction="wide")
#Change columns names 
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_LPG')] <- 'LPG'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Petroleum')] <- 'Petroleum'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Diesel')] <- 'Diesel'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Carbon')] <- 'Carbon'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Firewood')] <- 'Firewood'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Fuel_to_heat')] <- 'Fuel_to_heat'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Candles')] <- 'Candles'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Other')] <- 'Other'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_Electricity')] <- 'Electricity'
colnames(df_hsd_by_hsd_wd)[which(names(df_hsd_by_hsd_wd) == 'gasto_tri_NG')] <- 'NG'

df_hsd_by_hsd_wd
```

Store data of energy expenses by household
```{r}
write.csv(df_hsd_by_hsd_wd, file = 'outputs/2018_enigh_ener-exp_hsd_wide.csv',
          fileEncoding="UTF-8", row.names=FALSE)
```

## Descriptive Statistics
```{r}
descr(df_hsd_by_hsd_wd, stats = c("mean", "sd", "min", "med", "max", "n.valid", 
                                  "pct.valid"), transpose = TRUE)
```

### Box Plots
```{r plt-num-bulbs-box, fig.cap="Boxplots of energy expenses per household", results="hold"}
pal_plot <- c('grey','red', 'rgb(7,40,89)', 'green')
pal_plot <- setNames(pal_plot, c("elect", "ng", "candles", "lpg"))

#df_hsd_by_hsd_wd[is.na(df_hsd_by_hsd_wd)] = 0

pl_energy_exp_box <- plot_ly(type = 'box') %>%
  add_boxplot(y = df_hsd_by_hsd_wd$Electricity, 
              boxpoints = 'outliers', 
              name = "Electricity", 
              color = list(color =pal_plot['elect']),
              marker = list(color = pal_plot['elect']), 
              line=list(color = pal_plot['elect']) ) %>%
  add_boxplot(y = df_hsd_by_hsd_wd$NG, 
              boxpoints = 'outliers', 
              name = "Natural Gas", 
              color = list(color =pal_plot['ng']),
              marker = list(color = pal_plot['ng']), 
              line=list(color = pal_plot['ng']) ) %>%
  add_boxplot(y = df_hsd_by_hsd_wd$Candles, 
              boxpoints = 'outliers', 
              name = "Candles", 
              color = list(color =pal_plot['candles']),
              marker = list(color = pal_plot['candles']), 
              line=list(color = pal_plot['candles']) ) %>%
  add_boxplot(y = df_hsd_by_hsd_wd$LPG, 
              boxpoints = 'outliers', 
              name = "LPG", 
              color = list(color =pal_plot['lpg']),
              marker = list(color = pal_plot['lpg']), 
              line=list(color = pal_plot['lpg']) ) %>%
  layout(title = "Boxplots of Expenses by Household per Type of Energy", 
         yaxis = list(title = "$[MXN]", range = c(0, 5000)))

pl_energy_exp_box

plotly_IMAGE(pl_energy_exp_box, format = "png", out_file = "./figs/box_energy_expenses.png" )

```
# Analysis of Expenses by Dwelling


Grouping data sets by dwelling id and expense id (clave)
```{r, results ='hide'}
df_hsd_energy_by_dwell <- df_hsd_energy %>%
  group_by(folioviv, clave) %>% replace(is.na(.), 0) %>% 
  summarise(gasto_tri = sum(gasto_tri))

df_psn_energy_by_dwell <- df_person_energy %>%
  group_by(folioviv, clave) %>% replace(is.na(.), 0) %>% 
  summarise(gasto_tri = sum(gasto_tri))

head(df_hsd_energy_by_dwell)
head(df_psn_energy_by_dwell)
```


Merging data sets by id_household using energy expenses related data
```{r}
df_energy_by_dwell <- merge(df_hsd_energy_by_dwell, df_psn_energy_by_dwell, 
                          by="folioviv", all.y = TRUE, all.x = TRUE) %>% 
  replace(is.na(.), 0)

```

Reshaping data
```{r}
df_hsd_energy_by_dwell <- as.data.frame(df_hsd_energy_by_dwell)

df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G009'] <- 'LPG'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G010'] <- 'Petroleum'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G011'] <- 'Diesel'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G012'] <- 'Carbon'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G013'] <- 'Firewood'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G014'] <- 'Fuel_to_heat'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G015'] <- 'Candles'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='G016'] <- 'Other'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='R001'] <- 'Electricity'
df_hsd_energy_by_dwell$clave[df_hsd_energy_by_dwell$clave=='R003'] <- 'NG'

#df_hsd_energy_by_dwell$clave <-as.factor(df_hsd_energy_by_dwell$clave)
df_hsd_by_dwell_wd <- reshape(df_hsd_energy_by_dwell, v.names = 'gasto_tri',
                            timevar='clave', idvar="folioviv", sep = "_",
                            direction="wide")
#Change columns names 
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_LPG')] <- 'LPG'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Petroleum')] <- 'Petroleum'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Diesel')] <- 'Diesel'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Carbon')] <- 'Carbon'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Firewood')] <- 'Firewood'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Fuel_to_heat')] <- 'Fuel_to_heat'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Candles')] <- 'Candles'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Other')] <- 'Other'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_Electricity')] <- 'Electricity'
colnames(df_hsd_by_dwell_wd)[which(names(df_hsd_by_dwell_wd) == 'gasto_tri_NG')] <- 'NG'

df_hsd_by_dwell_wd
```

Adding weighting factor linked to the representation of dwellings
```{r, results = 'hide'}
df_dwell_factors <- subset(df_dwell, select=c('folioviv', 'est_socio', 'tam_loc', 
                                              'est_dis', 'upm', 'factor'))

df_hsd_by_dwell_wd <- merge(df_hsd_by_dwell_wd, df_dwell_factors, 
                          by="folioviv", all.y = TRUE, all.x = TRUE)
head(df_hsd_by_dwell_wd)
```

## Descriptive Statistics
```{r}
descr(df_hsd_by_dwell_wd, stats = c("mean", "sd", "min", "med", "max", "n.valid", "pct.valid"),
      weights = df_hsd_by_dwell_wd$factor,
      transpose = TRUE)
```

### Box Plots
```{r plt-exp-dwell-box, fig.cap="Boxplots of energy expenses per dwelling", results="hold"}
pal_plot <- c('grey','red', 'rgb(7,40,89)', 'green')
pal_plot <- setNames(pal_plot, c("elect", "ng", "candles", "lpg"))

#df_hsd_by_dwell_wd[is.na(df_hsd_by_dwell_wd)] = 0

pl_energy_exp_box <- plot_ly(type = 'box') %>%
  add_boxplot(y = df_hsd_by_dwell_wd$Electricity, 
              boxpoints = 'outliers', 
              name = "Electricity", 
              color = list(color =pal_plot['elect']),
              marker = list(color = pal_plot['elect']), 
              line=list(color = pal_plot['elect']) ) %>%
  add_boxplot(y = df_hsd_by_dwell_wd$NG, 
              boxpoints = 'outliers', 
              name = "Natural Gas", 
              color = list(color =pal_plot['ng']),
              marker = list(color = pal_plot['ng']), 
              line=list(color = pal_plot['ng']) ) %>%
  add_boxplot(y = df_hsd_by_dwell_wd$Candles, 
              boxpoints = 'outliers', 
              name = "Candles", 
              color = list(color =pal_plot['candles']),
              marker = list(color = pal_plot['candles']), 
              line=list(color = pal_plot['candles']) ) %>%
  add_boxplot(y = df_hsd_by_dwell_wd$LPG, 
              boxpoints = 'outliers', 
              name = "LPG", 
              color = list(color =pal_plot['lpg']),
              marker = list(color = pal_plot['lpg']), 
              line=list(color = pal_plot['lpg']) ) %>%
  layout(title = "Boxplots of Expenses per  Dwelling by Type of Energy", 
         yaxis = list(title = "$[MXN]", range = c(0, 5000)))

pl_energy_exp_box

plotly_IMAGE(pl_energy_exp_box, format = "png", out_file = "./figs/box_energy_exp_byDwell.png" )

```
Store data of energy expenses by dwelling
```{r}
write.csv(df_hsd_by_dwell_wd, file = 'outputs/2018_enigh_ener-exp_dwell_wide.csv',
          fileEncoding="UTF-8", row.names=FALSE)
```

## Survey Design
```{r}
df_dwell$tam_loc<- factor(df_dwell$tam_loc,
                   levels = c(1, 2, 3, 4),
                   labels = c(">=100k", "15K-100k", "2.5K-15K", "<2.5K") )

df_dwell$est_socio<- factor(df_dwell$est_socio,
                   levels = c(1, 2, 3, 4),
                   labels = c("low", "med-low", "med-high", "high") )
```


```{r, results = 'hide'}
df_hsd_by_dwell_wd$upm <- NULL
df_hsd_by_dwell_wd$factor <- NULL
df_hsd_by_dwell_wd$est_dis <- NULL
df_hsd_by_dwell_wd$tam_loc <- NULL
df_hsd_by_dwell_wd$est_socio <- NULL

df_dwell_sub <- subset(df_dwell, select=c('folioviv', 'est_socio', 'tam_loc', 
                                                    'upm', 'factor', 'est_dis'))
head(df_dwell_sub)

df_energy_by_folio <- merge(df_dwell_sub, df_hsd_by_dwell_wd, 
                          by="folioviv", all.y = TRUE, all.x = TRUE) %>% 
  replace(is.na(.), 0)

head(df_energy_by_folio)
```

```{r, results='hide'}
#Using Survey Design Construction
svd_dwell <- svydesign(id=~upm, strata=~est_dis, 
                 data=df_energy_by_folio, weights=~factor)

#Information about the survey design

class(svd_dwell)

#View the number of unique PSUs (clusters) in this survey design, by referring 
# to the clu column from the original data.frame:
length(unique(df_energy_by_folio$upm))

#View the number of unique strata in this survey design, by referring to the 
# est_dis column from the original data.frame:
length(unique(df_energy_by_folio$est_dis))
```


```{r}
tb_mean_energy  <- svyby(~Electricity, ~est_socio + tam_loc, svd_dwell, svymean, keep.var=TRUE)
tb_mean_energy
```

tb.mean.ing.exp.soc.mun <- svyby(~ing_cor + gasto_mon, ~est_socio + tam_loc, svd.household, svymean, keep.var=TRUE)
#tb.mean.ing.exp.soc.mun

```{r}
# Convert it to a table
tb_mean_energy %>% 
  ftable() %>%
  round(3)

```

```{r}
df_mean_energy_bytamloc <- tb_mean_energy %>% 
  group_by(tam_loc) 
df_mean_energy_bytamloc
```

```{r, fig.height=6, dev='jpeg', fig.path = 'figs/'}
colnames(df_mean_energy_bytamloc)[which(names(df_mean_energy_bytamloc) == 'est_socio')] <- 'Socioeconomic_Status'

colnames(df_mean_energy_bytamloc)[which(names(df_mean_energy_bytamloc) == 'tam_loc')] <- 'Municipality_Size'


#tb.mean.ing.exp.soc.mun$est_socio
p <- ggplot(df_mean_energy_bytamloc, aes(x=Municipality_Size, y=Electricity, 
                                    ymin = Electricity-se, ymax = Electricity+se,
                                    fill=Socioeconomic_Status)) + 
  geom_bar(stat = "identity", position=position_dodge()) + 
  geom_errorbar(position=position_dodge()) + 
  scale_fill_brewer(palette="Greens") +
  scale_y_continuous(name="Electricity Expenses [MXN]", labels=dollar) +
  ggtitle("Quarterly Expense in Electricity per Dwelling")

p <- p + theme(plot.title = element_text(color="black", size=16, face="bold"))
p

#ggsave("mtcars.pdf", width = 4, height = 4)
ggsave("./figs/ElectricityExpenses.png", width = 8, height = 6, units = "in")
```